BIP-Predictor

### 2025-05-13 Sweave/BIP_predictor.Rnw
up

# Select M-SSA components: use M-SSA output tracking BIP
sel_vec_mssa_comp<-select_vec_multi
sel_vec_mssa_comp<-c("BIP")
# Select indicators for direct forecast: all indicators lead to overfitting (worse out-of-sample performances)
# Best design out-of-sample is based on ifo_c and ESI
sel_vec_direct_forecast<-select_vec_multi
sel_vec_direct_forecast<-c("ifo_c","ESI")
# M-SSA looses L observations at start when compared to direct forecast or mean benchmarks: 
#   -therefore the samples are slightly different
# If align_sample==T then the first L observations of mean and direct forecasts are removed to align with M-SSA
#   -Avoid random differences due to unequal samples.
align_sample<-T
# Shift BIP forward by shift_vec + publication lag
shift_vec<-0:5
if (recompute_results)
{
# Select indicators for direct forecast: use all indicators
  sel_indicator_out_sample<-select_vec_multi
# Initialize performance matrices
  MSE_oos_mssa_comp_without_covid_mat<-MSE_oos_mssa_comp_mat<-p_mat_mssa<-p_mat_mssa_components<-p_mat_mssa_components_without_covid<-p_mat_direct<-rRMSE_mSSA_comp_direct<-rRMSE_mSSA_comp_mean<-rRMSE_mSSA_comp_direct_without_covid<-rRMSE_mSSA_comp_mean_without_covid<-rRMSE_mSSA_direct_mean_without_covid<-rRMSE_mSSA_direct_mean<-p_mat_direct_without_covid<-matrix(nrow=length(shift_vec),ncol=length(h_vec))
  final_components_preditor_array<-oos_components_preditor_array<-array(dim=c(length(shift_vec),length(h_vec),nrow(x_mat)))
  dimnames(final_components_preditor_array)<-dimnames(oos_components_preditor_array)<-list(paste("shift=",shift_vec,sep=""), paste("h=",h_vec,sep=""),rownames(x_mat))

@@

# Apply the previous function: compute GARCH, WLS regression, out-of-sample MSEs and p-values    
      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec,align_sample)
# Retrieve out-of-sample performances 
# a. p-values with/without Pandemic    
      p_mat_mssa_components[shift+1,k]<-perf_obj$p_value
      p_mat_mssa_components_without_covid[shift+1,k]<-perf_obj$p_value_without_covid
# b. MSE forecast error out-of-sample
#   -M-SSA components with/without Pandemic    
      MSE_oos_mssa_comp_mat[shift+1,k]<-MSE_oos_mssa_comp<-perf_obj$MSE_oos
      MSE_oos_mssa_comp_without_covid_mat[shift+1,k]<-MSE_oos_mssa_comp_without_covid<-perf_obj$MSE_oos_without_covid
#   -mean-benchmark with/without Pandemic    
      MSE_mean_oos<-perf_obj$MSE_mean_oos
      MSE_mean_oos_without_covid<-perf_obj$MSE_mean_oos_without_covid
# Here we retrieve the final in-sample predictor (based on the full-sample WLS regression) as well as the 
#    real-time out-of-sample predictor (re-adjusted to new data at each time point)
# We can plot both predictors to illustrate revisions (due to WLS estimation at each time point), see below
      final_in_sample_preditor<-perf_obj$final_in_sample_preditor
      final_components_preditor_array[shift+1,j+1,(nrow(x_mat)-length(final_in_sample_preditor)+1):nrow(x_mat)]<-final_in_sample_preditor
      cal_oos_pred<-perf_obj$cal_oos_pred
      oos_components_preditor_array[shift+1,j+1,(nrow(x_mat)-length(cal_oos_pred)+1):nrow(x_mat)]<-cal_oos_pred
# We can also obtain the regression weights to track changes (systematic vs. noisy revisions) over time
# Note: the variable will be overwritten, i.e., we keep only the last run through the double loop, 
#   corresponding to maximal shift and maximal forecast horizon, see exercise 2.2 below 
      track_weights<-perf_obj$track_weights
      track_weights_array[shift+1,j+1,(nrow(x_mat)-nrow(track_weights)+1):nrow(x_mat),]<-track_weights

#----------------      
# B. Direct forecasts
# -The main difference to M-SSA above is the specification of the explanatory variables in the data 
#     matrix dat: we here use x_mat instead of mssa_array. 
#   -We select all indicators (one could easily change this setting but results are only marginally effected as long as ifo and ESi are included)
#   -Note that the data matrix here does not depend on j, in contrast  to the M-SSA components above    
      dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),x_mat[,sel_vec_direct_forecast])
      rownames(dat)<-rownames(x_mat)
# Two variants for data set:  
      if (T)
      {
# 1. Same sample as M-SSA (due to filter initialization the first L observations are lost) 
        dat<-dat[L:nrow(dat),]
# Same length as M-SSA      
        nrow(na.exclude(dat))
      } else
      {
# 2. Full data, including the first L observations (makes sense since the data is available) 
        dat<-dat[1:nrow(dat),]
      }
      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec,align_sample)
# Retrieve out-of-sample performances: p-values and forecast MSE, with/without Pandemic 
      p_mat_direct[shift+1,k]<-perf_obj$p_value 
      p_mat_direct_without_covid[shift+1,k]<-perf_obj$p_value_without_covid 
      MSE_oos_direct<-perf_obj$MSE_oos
      MSE_oos_direct_without_covid<-perf_obj$MSE_oos_without_covid

>>>>

# Select M-SSA components: use M-SSA output tracking BIP
sel_vec_mssa_comp<-select_vec_multi
sel_vec_mssa_comp<-c("BIP")
# Select indicators for direct forecast: all indicators lead to overfitting (worse out-of-sample performances)
# Best design out-of-sample is based on ifo_c and ESI
sel_vec_direct_forecast<-select_vec_multi
sel_vec_direct_forecast<-c("ifo_c","ESI")
# M-SSA looses L observations at start when compared to direct forecast or mean benchmarks: 
#   -therefore the samples are slightly different
# If align_sample==T then the first L observations of mean and direct forecasts are removed to align with M-SSA
#   -Avoid random differences due to unequal samples.
align_sample<-T
reg_type<-"OLS"
reg_type<-"LASSO"
reg_type<-"Ridge"
lambda_reg<-10
# Shift BIP forward by shift_vec + publication lag
shift_vec<-0:5
if (recompute_results)
{
# Select indicators for direct forecast: use all indicators
  sel_indicator_out_sample<-select_vec_multi
# Initialize performance matrices
  MSE_oos_mssa_comp_without_covid_mat<-MSE_oos_mssa_comp_mat<-p_mat_mssa<-p_mat_mssa_components<-p_mat_mssa_components_without_covid<-p_mat_direct<-rRMSE_mSSA_comp_direct<-rRMSE_mSSA_comp_mean<-rRMSE_mSSA_comp_direct_without_covid<-rRMSE_mSSA_comp_mean_without_covid<-rRMSE_mSSA_direct_mean_without_covid<-rRMSE_mSSA_direct_mean<-p_mat_direct_without_covid<-matrix(nrow=length(shift_vec),ncol=length(h_vec))
  final_components_preditor_array<-oos_components_preditor_array<-array(dim=c(length(shift_vec),length(h_vec),nrow(x_mat)))
  dimnames(final_components_preditor_array)<-dimnames(oos_components_preditor_array)<-list(paste("shift=",shift_vec,sep=""), paste("h=",h_vec,sep=""),rownames(x_mat))

@@

# Apply the previous function: compute GARCH, WLS regression, out-of-sample MSEs and p-values    
      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec,align_sample,reg_type,lambda_reg)
# Retrieve out-of-sample performances 
# a. p-values with/without Pandemic    
      p_mat_mssa_components[shift+1,k]<-perf_obj$p_value
      p_mat_mssa_components_without_covid[shift+1,k]<-perf_obj$p_value_without_covid
# b. MSE forecast error out-of-sample
#   -M-SSA components with/without Pandemic    
      MSE_oos_mssa_comp_mat[shift+1,k]<-MSE_oos_mssa_comp<-perf_obj$MSE_oos
      MSE_oos_mssa_comp_without_covid_mat[shift+1,k]<-MSE_oos_mssa_comp_without_covid<-perf_obj$MSE_oos_without_covid
#   -mean-benchmark with/without Pandemic    
      MSE_mean_oos<-perf_obj$MSE_mean_oos
      MSE_mean_oos_without_covid<-perf_obj$MSE_mean_oos_without_covid
# Here we retrieve the final in-sample predictor (based on the full-sample WLS regression) as well as the 
#    real-time out-of-sample predictor (re-adjusted to new data at each time point)
# We can plot both predictors to illustrate revisions (due to WLS estimation at each time point), see below
      final_in_sample_preditor<-perf_obj$final_in_sample_preditor
      final_components_preditor_array[shift+1,j+1,(nrow(x_mat)-length(final_in_sample_preditor)+1):nrow(x_mat)]<-final_in_sample_preditor
      cal_oos_pred<-perf_obj$cal_oos_pred
      oos_components_preditor_array[shift+1,j+1,(nrow(x_mat)-length(cal_oos_pred)+1):nrow(x_mat)]<-cal_oos_pred
# We can also obtain the regression weights to track changes (systematic vs. noisy revisions) over time
# Note: the variable will be overwritten, i.e., we keep only the last run through the double loop, 
#   corresponding to maximal shift and maximal forecast horizon, see exercise 2.2 below 
      track_weights<-perf_obj$track_weights
      track_weights_array[shift+1,j+1,(nrow(x_mat)-nrow(track_weights)+1):nrow(x_mat),]<-track_weights

#----------------      
# B. Direct forecasts
# -The main difference to M-SSA above is the specification of the explanatory variables in the data 
#     matrix dat: we here use x_mat instead of mssa_array. 
#   -We select all indicators (one could easily change this setting but results are only marginally effected as long as ifo and ESi are included)
#   -Note that the data matrix here does not depend on j, in contrast  to the M-SSA components above    
      dat<-cbind(c(x_mat[(shift+lag_vec[1]+1):nrow(x_mat),1],rep(NA,shift+lag_vec[1])),x_mat[,sel_vec_direct_forecast])
      rownames(dat)<-rownames(x_mat)
# Two variants for data set:  
      if (T)
      {
# 1. Same sample as M-SSA (due to filter initialization the first L observations are lost) 
        dat<-dat[L:nrow(dat),]
# Same length as M-SSA      
        nrow(na.exclude(dat))
      } else
      {
# 2. Full data, including the first L observations (makes sense since the data is available) 
        dat<-dat[1:nrow(dat),]
      }
      perf_obj<-optimal_weight_predictor_func(dat,in_out_separator,use_garch,shift,lag_vec,align_sample,reg_type,lambda_reg)
# Retrieve out-of-sample performances: p-values and forecast MSE, with/without Pandemic 
      p_mat_direct[shift+1,k]<-perf_obj$p_value 
      p_mat_direct_without_covid[shift+1,k]<-perf_obj$p_value_without_covid 
      MSE_oos_direct<-perf_obj$MSE_oos
      MSE_oos_direct_without_covid<-perf_obj$MSE_oos_without_covid
